Centers and centralizers of higher enveloping algebras of Lie algebras

### 2020-09-09 draft.tex
added large N section

\owen{Point out how to use LQT map and deduce existence of $\EE_n$ algebra structure on a shifted copy of cyclic cohomology of~$\Omega^*_c$. Also do the Loday versions that use the other infinite series.}

\owen{The weird and interesting bit is to compute the factorization homology over a nontrivial (closed) manifold.}

>>>>

\owen{Please check this carefully, Greg. The lemma is meaningful even if the proposition is bogus.}

In the case of the infinite series of Lie algebras, 
there are theorems that tell us about the limiting behavior of Lie algebra chains and cochains as $N$ goes to infinity.
These can be applied to our models of the center to obtain novel $\EE_n$ algebras.

Let $\Cyc_*(A)$ denote the {\em cyclic chains} of a dg associative algebra~$A$.

\def\gl{\frak{gl}}

\begin{prp}
There is a quasi-isomorphism of $\EE_n$ algebras
\[
{\frak Z}_n(\frak{gl}_\infty) \xto{\simeq} \clies(\frak{\gl}_\infty) \otimes \Sym(\Cyc_*(k[-n])[1])  ,
\]
and so, in particular, the $N \to \infty$ limit of ${\frak Z}_n(\frak{gl}_N)$ is commutative.
\end{prp}

\begin{proof}
The Loday-Quillen-Tsygan (LQT) theorem --- for nonunital algebras --- provides a quasi-isomorphism
\[
\cliels(\gl_\infty(A))^{\gl_\infty(k)} \xto{\simeq} \Sym(\Cyc_*(A)[1])
\] 
and it is functorial in the algebra~$A$.
\owen{I used {\em invariants} here rather than {\em coinvariants}. Is that allowed?}
Now
\[
\clies(\gl_\infty, \cliels(\gl_\infty(A)) \simeq \cliels(\gl_\infty) \otimes \cliels(\gl_\infty(A))^{\gl_\infty(k)},
\]
as shown in the proof of the LQT theorem.

We take $A = \Omega^*_c(V)$ and use the functoriality to see that we get a quasi-isomorphism of prefactorization algebras, as claimed.
\end{proof}

There are immediate analogs of this theorem where we replace $\gl_N$ by $\frak{so}_N$ or $\frak{sp}_N$ and where we replace cyclic chains by dihedral chains.

A striking consequence of the LQT map appears at the level of factorization homology.

\begin{lmm}
For a closed $n$-dimensional manifold~$M$, 
\begin{align*}
\int_M {\frak Z}_n(\frak{gl}_\infty) 
&\simeq \clies(\Omega^*(M) \otimes \gl_\infty, \Sym(\Cyc_*(\Omega^*(M))[1])) \\
&\simeq \clies(\Omega^*(M) \otimes \gl_\infty, \Sym(\Omega^*(LM/^h S^1))[1]))
\end{align*}
where $\Omega^*(LM/^h S^1)$ denotes some cochain model for the cochains of the homotopy quotient of the free loop space under the canonical rotation action on loops.
\end{lmm}

\owen{That phrasing could be improved.}

Implicit in the statement is that one must describe the needed (and non-obvious) action of $\gl_\infty(\Omega^*(M))$ on a symmetric algebra generated by cyclic chains, respectively the rotation-equivariant cochains on the free loop space~$LM$.

What makes this result striking is that it shows how the {\em cyclic loop space} $LM/^h S^1$ appears out of taking a large $N$ limit of factorization homology.
This kind of result is especially interesting from the perspective of physics, 
as a kind of topological gauge/string duality.
In Section~\ref{sec: TFT} we explain how the center $\frak{Z}_n(\gl_\infty)$ can be understood as the {\em quantum observables} of topological BF theory for the Lie algebra $\gl_\infty$.
In this sense, the left hand side of the lemma $\int_M {\frak Z}_n(\frak{gl}_\infty)$ describes the global quantum observables over the manifold $M$.
On the right hand side, we see the cyclic loop space of $M$,
which arises naturally as the configuration space of a closed string theory with target~$M$.

\begin{proof}
As $\Omega^*_c(M) = \Omega^*(M)$ because $M$ is closed, 
one applies the LQT theorem for {\em unital} algebras
\[
\cliels(\gl_\infty(\Omega^*(M))) \xto{\simeq} \Sym(\Cyc_*(\Omega^*(M))[1])
\]
to obtain the first weak equivalence.
(In fact, one must transfer the $\gl_\infty(\Omega^*(M))$ along the LQT map.)
Then Jones' theorem \cite{} provides the second weak equivalence.
\end{proof}
